Virtualisierung eines Echtzeit-Betriebssystems zur Steuerung eines Roboters mit Schwerpunkt auf die Einhaltung der Echtzeit

### 2024-05-16 Masterarbeit_Robotik_Pamuk.tex
compare kernels
qemu 2cpus
Hardware and OS configuration checklist
links

\section{Task priorities}
Es gibt grundsätzlich 4 Gruppen

\begin{table}[ht]
	\centering
	\caption{Overview of the priority groups and their relationships}
	\label{tab:priorities}
	\begin{tabular}{|c|c|}
		\hline
		\textbf{Prioritätsgruppe}    & \textbf{Bereich} \\ \hline
		Xenomai Priorität            & 0 bis 99         \\ \hline
		Linux RT Priorität           & 1 bis 99         \\ \hline
		Linux (Nice Level) Priorität & -20 bis 19       \\ \hline
		RTK Priorität                & 0 bis 14         \\ \hline
	\end{tabular}
\end{table}
\clearpage

\section{Memory Management}
Es gibt verschiedene Speicherbereiche

@@

Figure \ref{fig:max_latency_hardware} shows latency of hardware Salamander4.
\begin{figure}[H]
	\centering
	\includegraphics[width=0.8\columnwidth]{img/max_latency_hardware.png}
	\caption[Latency hardware]{Latency hardware}
	\label{fig:max_latency_hardware}
\end{figure}
\clearpage

\section{Salamander 4 Virtualisation}
In addition to providing Salamander 4 on its own hardware, SIGMATEK has also developed a virtualised version of this operating system. It was developped using Yocto, an open source project that allows customised Linux distributions to be created for embedded systems \cite{WelcomeYoctoProject}. The virtualisation runs in a QEMU environment, which is an open source tool for hardware virtualisation \cite{QEMU}. With the help of the script depicted in code \ref{script:qemu_def}, Salamander 4 is started together with the necessary hardware components in the QEMU environment. This makes it possible to run Salamander 4 on a variety of host systems, regardless of the specific hardware of the host. Upon generating the necessary files, Yocto generates a QEMU folder with the following components shown in code \ref{lst:ls}.

@@

\subsection{Generic Ubuntu}

\subsection{Real-Time Ubuntu}

@@

\clearpage
\section{KVM exit reasons}\label{sec:kvm_exit_reasons}

\subsection{APIC\_WRITE}

@@

\clearpage
\subsection{HLT}
\subsubsection{Description}
\subsubsection{Occurence Reduction}
\clearpage
\subsection{EPT\_MISCONFIG}
\subsubsection{Description}
\subsubsection{Occurence Reduction}
\clearpage
\subsection{PREEMPTION\_TIMER}
\clearpage
\subsection{EXTERNAL\_INTERRUPT}
\clearpage
\subsection{IO\_INSTRUCTION}
\clearpage
\subsection{EOI\_INDUCED}
\clearpage
\subsection{EPT\_VIOLATION}
\clearpage
\subsection{PAUSE\_INSTRUCTION}
\clearpage
\subsection{CPUID}
\clearpage
\subsection{MSR\_READ}
\clearpage

\chapter{Real-Time Performance Optimizations and Latency Reduction}\label{cha:latency_reduction}
\section{Host OS Optimization}

\subsection{Tasks and Events}

\subsection{CPU governor}

@@

Figure \ref{fig:gnuplot_max_max_latency_default}
\begin{figure}[H]
	\centering
	\includegraphics[width=0.8\columnwidth]{masterthesis-documentation/docs/sigmatek/xenomai/default/gnuplot_max_latency_default.png}
	\caption[gnuplot latency no taskset]{gnuplot latency no taskset}
	\label{fig:gnuplot_max_max_latency_default}
\end{figure}

Figure \ref{fig:gnuplot_max_latency_with_taskset}
\begin{figure}[H]
	\centering
	\includegraphics[width=0.8\columnwidth]{masterthesis-documentation/docs/sigmatek/xenomai/with_taskset/gnuplot_max_latency_with_taskset.png}
	\caption[gnuplot latency with taskset]{gnuplot latency with taskset}
	\label{fig:gnuplot_max_latency_with_taskset}
\end{figure}

@@

\subsection{CPU isolation}
Isolating CPUs involves removing all user-space threads and unbound kernel threads since bound kernel threads are tied to specific CPUs and hence cannot be moved. Also, modifying the \texttt{proc/irq/IRQ\_NUMBER/smp\_affinity} property of each Interrupt IRQ\_NUMBER in the system is part of this process, as described later in section \ref{sec:irq_handling}. Output \ref{output:pse} shows the user and kernel tasks that run on CPU 19. After the isolation, user tasks other than the QEMU process have been removed from running on this CPU. Only few critical kernel threads that are tied to this CPU still take CPU time.

@@

\subsection{Interrupt Requests Handling}\label{sec:irq_handling}
Once the CPUs were isolated, Interrupt Requests Handling was the next step. Interrupt Requests are used to send a signal to the CPU, prompting it to 'interrupt' its current task and divert its attention to another task. This allows hardware devices to communicate with the CPU through frequent context switches, which can lead to performance degradation, especially in high-performance computing or real-time scenarios. To mitigate this, the IRQs needed to be removed from the isolated CPUs. This was done by manipulating a file in the proc filesystem, namely \texttt{/proc/irq/<IRQ>/smp\_affinity}. The value in the \texttt{smp\_affinity} file is a bitmask in hexadecimal format. Each bit in this mask corresponds to a CPU in the system. The least significant bit (LSB) on the right corresponds to the first CPU (CPU0), and the significance increases towards the left until CPU19. In a system with 20 CPUs, if every CPU was reserved for one IRQ, the value for \texttt{smp\_affinity} would be FFFFF. The script in code \ref{script:smp_affinity} was written to check and log the distribution of Interrupt Requests across each CPU in Salamander 4.

@@

\noindent By changing the values of the \texttt{smp\_affinity} files of the respective IRQs, the assignment of IRQs was controlled so that they would not be handled by the isolated CPU. This reduced the interruptions caused by IRQs and the isolated CPUs were able to focus more on their assigned tasks.

\subsection{Kernel tuning}

\section{Guest OS Optimization}

\subsection{Tasks and Events}

@@

    exec  qemu-system-x86_64 -M pc,accel=kvm -kernel ./bzImage \
    -m 2048 -drive file=salamander-image-sigmatek-core2.ext4,format=raw,media=disk \
    -append "console=ttyS0 console=tty1 root=/dev/sda rw panic=1 sigmatek_lrt.QEMU=1 ip=dhcp rootfstype=ext4 schedstats=enable" \
    -net nic,model=e1000,netdev=e1000 -netdev bridge,id=e1000,br=nm-bridge \
    -fsdev local,security_model=none,id=fsdev0,path=drive-c -device virtio-9p-pci,id=fs0,fsdev=fsdev0,mount_tag=/mnt/drive-C \
    -device vhost-vsock-pci,guest-cid=3,id=vsock0 \
    -drive if=pflash,format=qcow2,file=ovmf.code.qcow2 \
    -no-reboot -nographic
\end{lstlisting}
\end{minipage}

@@

\cite{pixelartSIGMATEKKompletteAutomatisierungssysteme}
\cite{Tracecmd}
\cite{KernelShark}
\cite{XenomaiXenomai}
\cite{WelcomeYoctoProject}
\cite{QEMU}
\cite{linPerformanceEvaluationXenomai}
\cite{abediRTNFPredictableLatency2019}
\cite{broskyShieldedProcessorsGuaranteeing2003}
\cite{casiniLatencyAnalysisVirtualization2021}
\cite{cinqueVirtualizingMixedCriticalitySystems2022}
\cite{garcia-vallsChallengesRealtimeVirtualization2014}
\cite{guStateoftheArtSurveyRealTime2012}
\cite{HardRealTime}
\cite{javierperezHowRealTime2022}
\cite{kirovaImpactModernVirtualization2019}
\cite{kiszkaLinuxRealTimeHypervisor}
\cite{lutsykPipelinedMulticoreMachine2020}
\cite{malallahComprehensiveStudyKernel2021}
\cite{masrurVMBasedRealTimeServices2010}
\cite{mckenneyRealTimeVs}
\cite{pielAsymmetricSchedulingLoad2006}
\cite{queirozTestingLimitsGeneralpurpose2023}
\cite{reghenzaniRealTimeLinuxKernel2020}
\cite{yoonRealTimePerformanceAnalysis}
\clearpage

@@

The Linux kernel with the Dovetail interface, which allows the kernel to handle real-time tasks with low latency.
The Cobalt real-time core, which is responsible for scheduling and executing real-time tasks.
In Xenomai 4, you need:
The Linux kernel with the EVL interface, which, like Dovetail, enables the kernel to handle real-time tasks with low latency.
The libevl library, which provides a user interface to the EVL core, allowing applications to use the real-time capabilities provided by EVL.
So, while both versions require a Linux kernel with a real-time interface (Dovetail for Xenomai 3, EVL for Xenomai 4), Xenomai 3 uses a real-time core (Cobalt), while Xenomai 4 uses a library (libevl) to provide applications with access to its real-time capabilities.
\end{comment}

\clearpage
\chapter{Discussion}\label{cha:discussion}
\clearpage
\chapter{Summary and Outlook}\label{cha:summary_and_outlook}

@@

\phantomsection
\addcontentsline{toc}{chapter}{\listacroname}
\chapter*{\listacroname}
\begin{acronym}[XXXXX]
	\acro{CPU}[CPU]{Central Processing Unit}
	\acro{QEMU}[QEMU]{Quick Emulator}
	\acro{IRQ}[IRQ]{Interrupt Request}
\end{acronym}

>>>>

\clearpage
\section{Task priorities}
Es gibt grundsätzlich 4 Gruppen

\begin{table}[ht]
	\centering
	\caption{Overview of the priority groups and their relationships}
	\label{tab:priorities}
	\begin{tabular}{|c|c|}
		\hline
		\textbf{Prioritätsgruppe}    & \textbf{Bereich} \\ \hline
		Xenomai Priorität            & 0 bis 99         \\ \hline
		Linux RT Priorität           & 1 bis 99         \\ \hline
		Linux (Nice Level) Priorität & -20 bis 19       \\ \hline
		RTK Priorität                & 0 bis 14         \\ \hline
	\end{tabular}
\end{table}
\clearpage
\section{Memory Management}
Es gibt verschiedene Speicherbereiche

@@

Figure \ref{fig:max_latency_hardware} shows latency of hardware Salamander4.
\begin{figure}[H]
	\centering
	\includegraphics[width=0.8\columnwidth]{img/max_latency_hardware.png}
	\caption[Latency hardware]{Latency hardware}
	\label{fig:max_latency_hardware}
\end{figure}
\clearpage
\section{Salamander 4 Virtualisation}
In addition to providing Salamander 4 on its own hardware, SIGMATEK has also developed a virtualised version of this operating system. It was developped using Yocto, an open source project that allows customised Linux distributions to be created for embedded systems \cite{WelcomeYoctoProject}. The virtualisation runs in a QEMU environment, which is an open source tool for hardware virtualisation \cite{QEMU}. With the help of the script depicted in code \ref{script:qemu_def}, Salamander 4 is started together with the necessary hardware components in the QEMU environment. This makes it possible to run Salamander 4 on a variety of host systems, regardless of the specific hardware of the host. Upon generating the necessary files, Yocto generates a QEMU folder with the following components shown in code \ref{lst:ls}.

@@

\subsection{Generic Ubuntu}
\clearpage
\subsection{Real-Time Ubuntu}

@@

\clearpage
\section{KVM exit reasons}\label{sec:kvm_exit_reasons}
\subsection{APIC\_WRITE}

@@

\clearpage
\subsection{HLT}

\clearpage
\subsection{EPT\_MISCONFIG}

\clearpage
\subsection{PREEMPTION\_TIMER}

\clearpage
\subsection{EXTERNAL\_INTERRUPT}

\clearpage
\subsection{IO\_INSTRUCTION}

\clearpage
\subsection{EOI\_INDUCED}

\clearpage
\subsection{EPT\_VIOLATION}

\clearpage
\subsection{PAUSE\_INSTRUCTION}

\clearpage
\subsection{CPUID}

\clearpage
\subsection{MSR\_READ}

\chapter{Real-Time Performance Optimizations and Latency Reduction}\label{cha:latency_reduction}
\section{Host OS Optimization}
\subsection{Tasks and Events}
\clearpage
\subsection{CPU governor}

@@

Figure \ref{fig:gnuplot_max_latency_default}
\begin{figure}[H]
	\centering
	\includegraphics[width=0.8\columnwidth]{masterthesis-documentation/docs/sigmatek/xenomai/default/gnuplot_max_latency_default.png}
	\caption[gnuplot latency no taskset]{gnuplot latency no taskset}
	\label{fig:gnuplot_max_latency_default}
\end{figure}

Figure \ref{fig:gnuplot_max_latency_taskset}
\begin{figure}[H]
	\centering
	\includegraphics[width=0.8\columnwidth]{masterthesis-documentation/docs/sigmatek/xenomai/taskset/gnuplot_max_latency_taskset.png}
	\caption[gnuplot latency with taskset]{gnuplot latency with taskset}
	\label{fig:gnuplot_max_latency_taskset}
\end{figure}

@@

\clearpage
\subsection{CPU isolation}
Isolating CPUs involves removing all user-space threads and unbound kernel threads since bound kernel threads are tied to specific CPUs and hence cannot be moved. Also, modifying the \texttt{proc/irq/IRQ\_NUMBER/smp\_affinity} property of each Interrupt IRQ\_NUMBER in the system is part of this process, as described later in section \ref{sec:irq_handling}. Output \ref{output:pse} shows the user and kernel tasks that run on CPU 19. After the isolation, user tasks other than the QEMU process have been removed from running on this CPU. Only few critical kernel threads that are tied to this CPU still take CPU time.

@@

\clearpage
\subsection{Interrupt Requests Handling}\label{sec:irq_handling}
Once the CPUs were isolated, Interrupt Requests Handling was the next step. Interrupt Requests are used to send a signal to the CPU, prompting it to 'interrupt' its current task and divert its attention to another task. This allows hardware devices to communicate with the CPU through frequent context switches, which can lead to performance degradation, especially in high-performance computing or real-time scenarios. To mitigate this, the IRQs needed to be removed from the isolated CPUs. This was done by manipulating a file in the proc filesystem, namely \texttt{/proc/irq/<IRQ>/smp\_affinity}. The value in the \texttt{smp\_affinity} file is a bitmask in hexadecimal format. Each bit in this mask corresponds to a CPU in the system. The least significant bit (LSB) on the right corresponds to the first CPU (CPU0), and the significance increases towards the left until CPU19. In a system with 20 CPUs, if every CPU was reserved for one IRQ, the value for \texttt{smp\_affinity} would be FFFFF. The script in code \ref{script:smp_affinity} was written to check and log the distribution of Interrupt Requests across each CPU in Salamander 4.

@@

\noindent By changing the values of the \texttt{smp\_affinity} files of the respective IRQs, the assignment of IRQs was controlled so that they would not be handled by the isolated CPU. This reduced the interruptions caused by IRQs and the isolated CPUs were able to focus more on their assigned tasks.
\clearpage
\subsection{Kernel tuning}
\clearpage

\section{Guest OS Optimization}
\clearpage
\subsection{Tasks and Events}
\clearpage

@@

    exec  qemu-system-x86_64 -M pc,accel=kvm -kernel ./bzImage \
    -m 2048 -drive file=salamander-image-sigmatek-core2.ext4,format=raw,media=disk \
    -append "console=ttyS0 console=tty1 root=/dev/sda rw panic=1 sigmatek_lrt.QEMU=1 ip=dhcp rootfstype=ext4 schedstats=enable" \
    -net nic,model=e1000,netdev=e1000 -netdev bridge,id=e1000,br=nm-bridge \
    -fsdev local,security_model=none,id=fsdev0,path=drive-c -device virtio-9p-pci,id=fs0,fsdev=fsdev0,mount_tag=/mnt/drive-C \
    -device vhost-vsock-pci,guest-cid=3,id=vsock0 \
    -drive if=pflash,format=qcow2,file=ovmf.code.qcow2 \
    -no-reboot -nographic
\end{lstlisting}
\end{minipage}
\clearpage

\chapter{Real-Time Robotic Application}\label{cha:real-time-testing}

\section{VARAN}
\clearpage
\section{Robotic Application}
\clearpage

@@

\cite{pixelartSIGMATEKKompletteAutomatisierungssysteme}
\cite{Tracecmd}
\cite{KernelShark}
\cite{XenomaiXenomai}
\cite{WelcomeYoctoProject}
\cite{QEMU}
\cite{linPerformanceEvaluationXenomai}
\cite{abediRTNFPredictableLatency2019}
\cite{broskyShieldedProcessorsGuaranteeing2003}
\cite{casiniLatencyAnalysisVirtualization2021}
\cite{cinqueVirtualizingMixedCriticalitySystems2022}
\cite{garcia-vallsChallengesRealtimeVirtualization2014}
\cite{guStateoftheArtSurveyRealTime2012}
\cite{HardRealTime}
\cite{javierperezHowRealTime2022}
\cite{kirovaImpactModernVirtualization2019}
\cite{kiszkaLinuxRealTimeHypervisor}
\cite{lutsykPipelinedMulticoreMachine2020}
\cite{malallahComprehensiveStudyKernel2021}
\cite{masrurVMBasedRealTimeServices2010}
\cite{mckenneyRealTimeVs}
\cite{pielAsymmetricSchedulingLoad2006}
\cite{queirozTestingLimitsGeneralpurpose2023}
\cite{reghenzaniRealTimeLinuxKernel2020}
\cite{yoonRealTimePerformanceAnalysis}
\cite{adamPerformanceAssessmentLinux2021}
\clearpage

@@

The Linux kernel with the Dovetail interface, which allows the kernel to handle real-time tasks with low latency.
The Cobalt real-time core, which is responsible for scheduling and executing real-time tasks.
In Xenomai 4, you need:
The Linux kernel with the EVL interface, which, like Dovetail, enables the kernel to handle real-time tasks with low latency.
The libevl library, which provides a user interface to the EVL core, allowing applications to use the real-time capabilities provided by EVL.
So, while both versions require a Linux kernel with a real-time interface (Dovetail for Xenomai 3, EVL for Xenomai 4), Xenomai 3 uses a real-time core (Cobalt), while Xenomai 4 uses a library (libevl) to provide applications with access to its real-time capabilities.
\end{comment}
\clearpage

\chapter{Discussion}\label{cha:discussion}
\clearpage

\chapter{Summary and Outlook}\label{cha:summary_and_outlook}

@@

\phantomsection
\addcontentsline{toc}{chapter}{\listacroname}

\chapter*{\listacroname}
\begin{acronym}[XXXXX]
	\acro{CPU}[CPU]{Central Processing Unit}
	\acro{QEMU}[QEMU]{Quick Emulator}
	\acro{IRQ}[IRQ]{Interrupt Request}
\end{acronym}
